---
title: "Who Represents You?"
output: 
  html_document:
    css: styles.css
---


```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F, echo = F)
source("civic-info.R")
source("propublica.R")
library("knitr")
```

##Your Representatives
This is a report about the political representatives based on the address entered (`r address`). This report draws upon data from the [ProPublica API](https://projects.propublica.org/api-docs/congress-api/) and the [Google Civic Data API](https://developers.google.com/civic-information/). As an overview, here is a summary of _who_ represents the address given (from the **local** to the **national** level).

```{r table, img}
kable(officials_kable)
```


##House of Representatives
Now let's take a look at **all** of the congressional representatives (in the House of Representatives, _not_ the Senate) for the state in which the address was given. We see this breakdown by gender and political party for those members of the House:

```{r plot}
gender_plot
party_plot
```

##`r name`
Let's learn a bit more about one representative, _`r name`_. The representative is `r age` years old, and you can find them on twitter at [`r twitter`](`r twitter_url`). Interestingly, of the last `r num_votes` congressional votes, the representative was in accordance with the majority `r vote_agree_percent`% of the time.